appendix.Rmd

### 2024-11-30 manuscript_version/revision/appendix.Rmd
rearranging folders and updating the manuscript

---
title             : "Moral Foundations of U.S. Political News Organizations"
shorttitle        : "MORAL NEWS"

author:
  - name          : "William E. Padfield"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "901 S. National Ave, Springfield, MO, 65897"
    email         : "wpadfield@gmail.com"
  - name          : "Erin M. Buchanan, Ph.D."
    affiliation   : "2"
  - name          : "Kayden N. Jordan, Ph.D."
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Missouri State University"
  - id            : "2"
    institution   : "Harrisburg University of Science and Technology"

authornote: |
  William Padfield completed this project as part of his thesis requirements for a master's degree candidate in Psychology at Missouri State University. Submitted to Meta-Psychology. Participate in open peer review by commenting through hypothes.is directly on this preprint. The full editorial process of all articles under review at Meta-Psychology can be found following this link: https://tinyurl.com/mp-submissions

abstract: |
  Partisan differences and diviseness have become an increasing hot topic in psychological research. Many theories have been proposed to explain these differences and divisions including Moral Foundations Theory. The current research seeks to use a linguistic measure of Moral Foundations, the Moral Foundations Dictionary (MFD), to test the theory in terms of predicted partisan differences. In Study 1, we explore the MFD-Moral Foundations Questionnaire's measurement overlap using a multi-trait multi-method model. These latent trait models were used to inform possible improvement of the MFD. Next, we examined two replication extensions of previous work exploring the differences in political articles on moral differences Through web scraping, we extracted articles from popular partisan news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Study 1, we gathered political articles from four sources. We were unable to reveal significant differences in moral endorsements, but we solidified the method to be employed in further research. In Study 2, we expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral endorsements were found. Together with past work, the results shed doubt on the validity of the MFD as a reliable measurement tool.

keywords          : "politics, morality, psycholinguistics"

bibliography      : ["./metapsych_bib.bib"] # add new references

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
bibliography: references.bib # delete this 
---

```{r load_packages, include = FALSE}
library(papaja)
library(MuMIn)
library(reshape)
library(nlme)
library(ggplot2)
library(expss)
library(MOTE)
library(quanteda)
library(quanteda.textstats)
library(tm)
library(ngram)
library(ggplot2)
library(reshape)
library(knitr)
library(performance)
library(dplyr)
library(rio)
library(mice)
library(semPlot)
library(lavaan)

# to install this package 
# tmp <- tempfile()
# system2("git", c("clone", "--recursive",
#                  shQuote("https://github.com/leslie-huang/r-corpus.git"), shQuote(tmp)))
# devtools::install(tmp)
library(corpus)

knitr::opts_chunk$set(cache = TRUE)
options(tinytex.verbose = TRUE)
percentmiss <- function(x){ sum(is.na(x)) / length(x) * 100 }
```

The field of moral psychology has a long history with modern moral
psychology beginning in the late 1960s with Lawrence Kohlberg's theory
of moral development [@Kohlberg1977]. Since then, Kohlberg's theory has
been highly criticized with numerous theories and hypotheses proposed to
replace it. One of these theories is Moral Foundations Theory (MFT)
proposed by Haidt and colleagues [@haidt2004; @Haidt2007]. While the
theory itself has been widely criticized and debated, the goal of the
current work is to tackle a related measurement issue, the Moral
Foundations Dictionary [MFD, @Graham2009]. @Graham2009 developed the
MFD - a lexicon-based linguistic measure of MFT - arguing that moral
concerns and the propositions of MFT could be measured and tested with
the words people use. Here we first review evidence and criticisms
against this assumption. We then provide our own test of the MFD using
partisan news articles.

## Moral Foundations Measurement

Haidt and Joseph [-\@haidt2004] and Haidt and Graham [-@Haidt2007]
formulated MFT as a method by which to capture the entirety of humans'
moral domain. They argued that older theories of moral psychology were
focused primarily on issues of justice, fairness, and caring, which are
individually focused foundations of morality that align with the beliefs
of political liberals. In other words, moral psychology ignored the
valid moral foundations of conservatives. MFT holds that people's moral
domain can be mapped by quantifying their endorsement of five moral
foundations: *harm/care*, *fairness/reciprocity*, *ingroup/loyalty*,
*authority/respect*, and *purity/sanctity* [@Haidt2007]. Thus, they
proposed that the addition of these two additional domains could
adequately capture the broader spectrum of morality.

Measurement of MFT has traditionally focused on two options: 1) the
Moral Foundations Questionnaire (MFQ) - a typical Likert-type survey
instrument that measures moral relevance of each foundation to an
individual and the moral judgment of various scenarios depicting each
foundation, and 2) the Moral Foundations Dictionary (MFD) which was
designed to capture language's role in moral reasoning and justification
as used in speech and text [@Graham2009; @graham2011]. The MFQ is a
30-item scale with 15 items on the moral relevance and judgment
scenarios. The 15 moral relevance items are equally divided among the
five foundations and examine how relevant a condition is to making a
moral judgment on a scale of 1 (*not at all relevant*) to 6 (*extremely
relevant*). These relevance items include examples such as: "Whether or
not someone used violence (*harm*)," "Whether or not someone was denied
his or her rights (*fairness*)," "Whether or not someone showed a lack
of loyalty (*ingroup*)," "Whether or not an action caused chaos or
disorder (*authority*)," and "Whether or not someone did something
disgusting (*purity*)." The moral judgments items are also equally
divided between the foundations and ask on a six-point scale how much
one agrees with each of the statements. These judgment items include:
"One of the worst things a person can do is hurt a defenseless animal
(*harm*)," "Justice is the most important requirement of a society
(*fairness*)," "I am proud of my country's history (*ingroup*)," "Men
and women each have different roles to play in society (*authority*),"
and "Chastity is an important and valuable virtue (*purity*)."

The internal consistency of the MFQ from @Graham2011 was $\alpha$ = .73
averaged across subscales with a range of $\alpha$ = .65-.84. Across six
studies, the MFQ was found to have an average Cronbach's alpha of .63
for *harm*, .64 for *fairness*, .56 for *ingroup*, .59 for *authority*,
and .71 for *purity* [@Federico2013; @Graham2009; @Graham2012;
@Weber2013]. Test-retest reliability was *r* = .68-.82 using a sample of
123 college students. Confirmatory factor analysis supported a
well-fitted five-factor model (*harm/care*, *fairness/reciprocity*,
*ingroup/loyalty*, *authority/respect*, and *purity/sanctity*) over two,
individual (*harm* and *fairness*) versus group (*ingroup*, *authority*,
and *purity*) foundations, or three, autonomy (*harm*, *fairness*),
community (*ingroup*, *authority*), and divinity (*purity*) ethics,
foundations factor model. The five-factor structure also fit for
non-Western samples, thus, providing evidence of the MFQ
generalizability. Convergent validity was supported with correlations on
other measures of morality [@Schwartz; @Graham2011].

The MFD is composed of 259 words, with around 50 words assigned to each
of the five foundations. @Graham2009 created a preliminary list of words
that they believed would be associated with the five foundations. Words
such as *war* and *peace* should indicate a greater concern with *harm*
foundation whereas words such as *homeland* and *terrorism* should
indicate a greater concern with the *ingroup* foundation. The other
foundation dictionaries include *equal* and *justice* (*fairness*),
*honor* and *protest* (*authority*), and *holy* and *sin* (*purity*). To
validate the word sets, @Graham2009 examined the frequency of MFD words
in liberal and conservative sermons using the Linguistic Inquiry and
Word Count [LIWC; @Pennebaker2007] computer program. They found liberal
ministers used *harm*, *fairness*, and *ingroup* words more often than
conservative ministers who used *authority* and *purity* words more
often. Although conservative ministers were expected to use more
*ingroup* words based on political ideology and previous research, an
examination of the way liberal ministers used *ingroup* words revealed a
tendency for the use of *ingroup* words to glorify rebellion and promote
independence (i.e., the opposite direction from *ingroup* definitions).
Effect sizes indicated relatively sizable difference between liberal and
conservative sermons with Cohen's *d* values ranging from 0.56 to 1.27.

In contrast to this approach, @Sagi2013 explored how moral words were
used paired with other co-occurring concepts using Latent Semantic
Analysis (LSA). They examined three different moral issues in different
contexts to piece out specific moral words and their collocates. First,
they looked at how moral words were used in relation to the World Trade
Center compared to the Empire State Building in the New York Times from
1987-2007. After 9-11, the number of moral words associated with the
World Trade Center increased, specifically *harm* words from the MFD.
Second, they considered the changes in how moral words were paired with
mosque used in blogs as a response to the debate of building a mosque
near Ground Zero following 9-11. They found words from the MFD were used
more often with mosque during the main debate and then the co-occurrence
decreased afterwards. Lastly, they examined moral language tied to the
abortion debate in Congress. Republicans used more moral language
overall; more specifically, Republicans tended to use more words
associated with the *purity* foundation; while Democrats used more words
associated with the *fairness* foundation.

## Critiques of Moral Foundations

MFT has received criticism on the grounds that its assumptions regarding
moral intuitions have little empirical basis. @Suhler2011 criticized the
content and taxonomy of the five foundations and question whether or not
the foundations are sufficiently distinct as to stand as their own
foundation. Likewise, @Gray2015 argues that the measurement of the moral
foundations confounds morality with other constructs (such as weirdness)
and the moral foundations lack statistical and conceptual distinction.
@Schein2015 argues that harm-based morality is the most parsimonious and
logical explanation for MFT findings.

Beyond the critiques of the theory itself, many have also provided
strong critiques of the measurement using MFD. Conducting two close
conceptual replication studies using religious sermons and six extension
studies using U.S. Congressional speeches, party platforms, State of the
Union addresses, media outlets, and political organization texts,
@Frimer2020 found that the liberal-conservative differences found in the
original study replicated in only 30% of cases and the effect sizes were
over 30 times smaller than reported in @Graham2009. In a meta-analysis,
the theorized differences were only found for authority and purity
(e.g., conservatives used more authority and purity words). Loyalty was
contrary to theoretical predictions (e.g., used more by liberals), and
no differences were found for harm and fairness. Additionally,
@Frimer2020 tested the theory under the most forgiving conditions using
not only the original MFD but also the revised MFD2 as well using log
transformed scores and raw scores. Even under the best circumstances
(log transformed MFD2 scores), only two of the foundations showed
theorized differences but with very small effects. Other studies have
also found conflicting findings which fail to support the basic
assertions of MFT. Exploring the political Twitterverse, @Sterling2018
found differences in the MFD depend on political sophistication. For
example, harm words were used more by liberals only if they had low
levels of political sophistication; at high levels of political
sophistication, conservatives used more harm words. While the MFD has
not been extensively used (at least in published research), studies
which have use the MFD to test partisan differences have found mixed
results in terms of supporting MFT [@Clifford2013; @Frimer2015;
@Sagi2014]. Like @Frimer2020, we attempt to give the MFD the best chance
for success by incorporating a weighting method to address the
skewedness and sparsity of the linguistic data.

The most successful attempts to validate a linguistic measure of moral
foundations have used more complex methods (compared to the original
dictionary-based approach). These attempts generally fall into two
categories: human annotations and semantic representations. @Hopp2021
represents the first approach where instead of relying on 'experts', a
large sample of lay people were crowdsourced to manually annotate
document for content relevant to each foundation. While
resource-intensive, this method better captures how people may encounter
and judge moral issues in everyday life leading to a more ecologically
valid dictionary. @Garten2016 represents the second categories which
instead relies on data-driven semantic analysis. Using shallow neural
net models like word2vec [@mikolov_efficient_2013], moral foundations
were measured using semantic similarity vectors rather than simple word
counts. Here, we try to augment the simpler approach to add more context
to see if the MFD liberal-conservative differences can be replicated.

First, we use the MFQ and MFD in a multi-trait multi-method approach to
determine if the dictionary can be improved and mapped into the
theoretical MFT traits. This study examined the measurement overlap
between these two sources, as well as the latent traits after
controlling for the overlap. Using the results from this study, we
examined two conceptual replications of @Frimer2020. We test the
liberal-conservative difference proposed by MFT using partisan media
content and the MFD. We first test differences in a general news corpus
compared liberal sources (National Public Radio [NPR] and the New York
Times [NYT]) to conservative sources (Fox News and Breitbart). Second,
we examined news about specific political events: Brett Kavanaugh's U.S.
Supreme Court nomination and the U.S. Government Shutdown in 2018-2019.
In the second study, the news sources were expanded to cover NYT, NPR,
Slate, Huffington Post, and Politico on the liberal side and Fox News,
Breitbart, The Rush Limbaugh Show, The Blaze, and Sean Hannity on the
conservative side. To address potential issues of measurement with the
MFD, the MFD was combined with the information produced in Study 1 and
weighted by valence to create weighted percentages to better specify
endorsement. Like @Frimer2020, results were judged based on the
direction of the liberal-conservative difference compared MFT
predictions and the size of the effect compared to past studies.

# Study 1

# Method

We approached this study with the intention to answer a method question.
That is, this portion of the current research was conducted in order to
solidify the best method by which to analyze political news text under
the MFT framework while also alleviating some of the aforementioned
potential dictionary issues with the MFD. In this set of studies, we
examined the MFD and MFQ within the same study. The first two
sub-studies were used to develop a (potentially) larger dictionary for
the MFD and collect data on both scales. The original dictionary was
developed conceptually, identifying words that *should* go with each
moral trait. In Study 1.1 and 1.2, we collect data to expand the
dictionary. The last sub-study examined the MFD and MFQ using
multi-trait multi-method (MTMM) analysis using guidelines from
@byrne2001 and @widaman1985.

MTMM analyses use latent variable modeling to represent the underlying
structure of a proposed phenomenon, while also accounting for the
measurement of that phenomenon. As shown in Figure \@ref(fig:fig-mtmm),
MTMM models are programmed to represent the traits that are measured
(left hand side circles) and the methods used to measure those traits
(right hand side circles). The squares in the middle represent the
actual observed data, which is the MFQ questions and MFD scores. In the
first model, *correlated traits correlated methods*, the latent
variables are allowed to correlate on the traits and methods side, but
not cross correlate from traits to methods. All models are then compared
to this model to examine fit, convergent, and divergent validity. In the
second model, *no traits correlated methods*, the trait side of the
model is deleted and only methods are considered. The model should
significantly degrade in fit, as the traits should be useful in
explaining model variance for convergent validity (i.e., you do not want
a model that suggests all the variances is due to measurement).

In the third model, *perfectly correlated traits freely correlated
methods*, the traits are added back to the model, but each correlation
is set to one. The fit is compared to model 1, indicating if divergent
validity of the traits has occurred. A significant degrade in fit should
be found to indicate that trait correlations are not perfect. Last,
model four *correlated traits uncorrelated methods*, the traits are
allowed to naturally correlate but the methods correlations are set to
zero. In comparison with model 1, the model should be roughly equal for
divergent validity. This result would indicate that the methods each
represent a different component to the measurement of the model, and do
not represent the same measurement twice.

Given the goal of improved measurement, we compared model 1 of the MTMM
steps across several dictionaries to determine the best measurement for
the dictionary. To assess model fit, we used the Root Mean Squared Error
of Approximation [RMSEA, @steiger1990], the Root Mean Squared Residual
[RMSR, @jöreskog1971], the Confirmatory Fit Index [CFI, @bentler1990],
the Tucker-Lewis Index [TLI, @bentler1980]. The RMSEA and RMSR should be
low for adequate model fit (\< .10) and the CFI and TLI should be high
for adequate model fit (\> .90). For model comparison, we used a
$\Delta$ CFI of greater than .01 to indicate a significant degradation
in fit [@hu1999]. We will additionally use the standardized loadings for
each of the traits to determine which model has the best measurement of
moral foundations (\> .30 desired).

For clarity, we programmed the following models:

1.  Model 1 *Correlated Traits Correlated Methods*
    a.  Original MFD Words
    b.  Original MFD Words + Study 1.1 + Study 1.2 Words
    c.  Original MFD Words + Study 1.1 Reduced + Study 1.2 Words
    d.  Determine if model suggests a feasible structure (fit indices
        adequate + loadings adequate)
    e.  Pick the best model to continue (using model comparison)
2.  Model 2 *No Traits Correlated Methods*
3.  Model 3 *Perfectly Correlated Traits Correlated Methods*
4.  Model 4 *Correlated Traits Uncorrelated Methods*

```{r fig-mtmm, fig.cap="Representation of the multi-trait multi-method model for the MFD and MFQ. Not all measured variables (squares) are shown for ease of reading. The left hand side represents the traits or foundation areas of MFT, while the right hand side represents the two measurement tools for the MFT. "}
knitr::include_graphics("mtmm_mfd.png")
```

## Study 1.1

In this study, participants were given a word association task elicit
additional semantically related word forms for each moral foundation.
This data was used to expand the MFD for each latent area to determine
if these additional words could improve scores and test measurement
relation to the MFQ in Study 1.3.

### Participants

```{r demos-mtmm-study-3.1, include = FALSE}
exp1data <- import("../../study1_data/study1.1_fixed_associations.csv")

# clear out completely columns we don't need
crapcolumns <- c(1, 2, 3, 5:14)
study1.1DF <- exp1data[ , -crapcolumns]

# accuracy
summary(study1.1DF) #Check Q15_1 and Q23 for in-range scores -WP
study1.1DF$Q23 <- factor(study1.1DF$Q23,
                         levels = 1:3,
                         labels = c("Democrat", "Republican", "Independent"))

# missing data 
missing <- apply(study1.1DF, 1, percentmiss)
table(missing)
# a few missing here and there
# do not want to replace entire MFQ, other missing is political orientation
# no replacement

# outliers on MFQ only 
mahal1 <- mahalanobis(study1.1DF[ , -c(1:6,22,23)], 
                    colMeans(study1.1DF[ , -c(1:6,22,23)], na.rm = TRUE),
                    cov(study1.1DF[ , -c(1:6,22,23)], use="pairwise.complete.obs"))
#mahal1
cutoff1 <- qchisq(1 - .001, ncol(study1.1DF[ , -c(1:6,22,23)]))
df1 <- ncol(study1.1DF[ , -c(1:6,22,23)])
summary1 <- summary(mahal1 < cutoff1) # nas will be excluded because they don't have the full scale 
noout1.1 <- subset(study1.1DF, mahal1 < cutoff1)

# assumptions
# correlations for MFQ
correlations1 <- cor(noout1.1[,-c(1:6,22,23)], use="pairwise.complete.obs")
correlations1
symnum(correlations1)

# making the random stuff
random1 <- rchisq(nrow(noout1.1), 7)
fake1 <- lm(random1 ~ ., data = noout1.1[ , -c(1:6, 22, 23)])
# normal
standardized1 <- rstudent(fake1)
hist(standardized1, breaks=15) # some skew

# linear
{qqnorm(standardized1) 
abline(0,1)}
# fairly linear

# homog + s
fitvalues1 <- scale(fake1$fitted.values)
{plot(fitvalues1, standardized1) 
abline(0,0)
abline(v = 0)}
# actually pretty good minus that one outlier

# create subscales 
noout1.1$harmMFQ <- apply(noout1.1[ , 7:9], 1, sum)
noout1.1$fairMFQ <- apply(noout1.1[ , 10:12], 1, sum)
noout1.1$ingroupMFQ <- apply(noout1.1[ , 13:15], 1, sum)
noout1.1$authorityMFQ <- apply(noout1.1[ , 16:18], 1, sum)
noout1.1$purityMFQ <- apply(noout1.1[ , 19:21], 1, sum)

party <- table(noout1.1$Q23, useNA = "ifany")
```

`r nrow(exp1data)` participants were assessed from a large Midwestern
university. Participants were given course credit for their introductory
psychology course for completing the study. `r unname(summary1[4])`
participants had missing data on the MFQ and were excluded from
analyses. The MFQ data was screened for multivariate outliers with
Mahalanobis distance as described in @Tabachnick2012, and
`r unname(summary1[2])` outliers were found using
$\chi^2_{p < .001}$(`r df1`) = `r apa_num(cutoff1)`. These participants
were excluded from further analyses, representing `r nrow(noout1.1)`
final participants. Participants were asked to denote their political
party, and `r printnum(party["Democrat"]/sum(party)*100,digits = 1)`%
indicated they were Democrats,
`r printnum(party["Republican"]/sum(party)*100,digits = 1)`% were
Republican, and
`r printnum(party["Independent"]/sum(party)*100,digits = 1)`% indicated
they were Independent. `r printnum(party[4]/sum(party)*100,digits = 1)`%
of values were missing for this question.

### Materials and Procedure

A complete example of the survey can be found online at
<https://osf.io/5uqfn>. First, participants were given a description of
associative memory as the relation between words that comes about
through many pairings in writing and speech. Next, the free association
task, similar to that used in @nelson2004 and @dedeyne2019 was described
to the participants as listing the "first word that pops into mind". The
participants were then given three example free association cues,
*lost*, *old*, and *article*. For each cue, participants were asked to
write all the words that come to mind. To elicit free association to the
moral foundation areas, participants were given the following
instructions:

"Moral Foundations Theory states that when making moral
judgments/decisions, the concerns people have can be divided into five
categories. Below are labels of each of these five categories. You will
then be asked to list words you think are associated with each of the
labels."

Each of the foundation pairs were listed together (i.e., *harm/care*,
*fairness/reciprocity,* etc.) with a space for participants to list
their free association concepts. After the free association task,
participants were then given the 15-items from the moral relevance
section of the MFQ as described in the introduction. Last, participants
were asked to denote their political orientation from 1 *conservative*
to 10 *liberal*, as well as which political party they associated with:
Democrat, Republican, and Independent. The survey was delivered through
Qualtrics, and participants were recruited through the online
participant management system for the university (SONA). Each
participant signed an online consent form at the beginning of the study
and was given participation credit at the end of the study.

### Results

All data was screening for inaccurate responses, as well as missing
data, as described in the participant section. The final MFQ data was
screened for assumptions of normality, linearity, and homoscedasticity.
The MFQ data was slightly skewed but otherwise linear and homoscedastic
after accounting for skew. No transform was applied. The sum of each
moral foundation area was calculated in order to determine which words
were linked to their respective moral foundation. The average scores
were: harm (*M* = `r printnum(mean(noout1.1$harmMFQ))`, *SD* =
`r printnum(sd(noout1.1$harmMFQ))`), fairness (*M* =
`r printnum(mean(noout1.1$fairMFQ))`, *SD* =
`r printnum(sd(noout1.1$fairMFQ))`), ingroup (*M* =
`r printnum(mean(noout1.1$ingroupMFQ))`, *SD* =
`r printnum(sd(noout1.1$ingroupMFQ))`), authority (*M* =
`r printnum(mean(noout1.1$authorityMFQ))`, *SD* =
`r printnum(sd(noout1.1$authorityMFQ))`), and purity (*M* =
`r printnum(mean(noout1.1$purityMFQ))`, *SD* =
`r printnum(sd(noout1.1$purityMFQ))`). Participants free association
responses were processed using the *tm* [@feinerer2008] and *corpus*
[@perry2024] libraries after manual correction of spelling errors. Each
set of answers was cleaned for punctuation, English stop words (e.g.,
*the*, *an*, *of*) were removed, and each word was stemmed using the
English library in *tm*. We did not combine related words in this
section (i.e., *injure* and *injury*, which have different stems *injur*
and *injuri*) to allow for maximum coverage of different word forms
present in the dictionary. Additionally, with the use of automated
stemmers like that present in the *tm* library, leaving both word forms
in the dictionary would capture more of the concepts present in future
analyses with a different corpus without the requirement on the coder to
manually recode all word forms. Frequency counts of the stemmed words
were tabulated and only words mentioned with at least one percent
frequency were used in the subsequent analyses. The complete set of word
frequencies for each foundation can be found in our supplemental
materials.

For clarity, the processing pipeline for each separate domain was:

1.  Manual spell checking of free association responses
2.  Word tokenization and stemming
3.  Create a term by document matrix
4.  Remove punctuation and stopwords via the *tm* stopword list
5.  Calculate percent frequency across documents (all words saved
    online)
6.  Remove words with less than 1% mentions across participants

```{r harm-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q27 <- as.character(noout1.1$Q27)

# stem the data
for (i in 1:nrow(noout1.1)) {
  noout1.1$Q27[i] <- paste(unlist(
    text_tokens(noout1.1$Q27[i], stemmer = "en")), collapse = " ")
}

# create a corpus
harm_corpus <- Corpus(VectorSource(noout1.1$Q27))
harm_TDM <- as.matrix(TermDocumentMatrix(harm_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

# calculate the most frequent words
harm_freq <- data.frame(Word = rownames(harm_TDM),
                       Freq = rowSums(harm_TDM),
                       row.names = NULL)
harm_freq$Word <- as.character(harm_freq$Word)
harm_freq$percent <- harm_freq$Freq/nrow(noout1.1) *100

write.csv(harm_freq, "../../study1_data/study1.1_harm_freq.csv", row.names = F)

# subset out to only words with 1% mentions
harm_words <- harm_freq$Word[harm_freq$percent >=1]
# turn sideways to add to dataset
harm_TDM  <- as.data.frame(t(harm_TDM))
harm_TDM  <- harm_TDM[ , harm_words]

# final dataset for harm
harm_final <- cbind(noout1.1[ , c("ResponseId", "Q15_1", "Q23", "harmMFQ")],
                   harm_TDM)

harm_cor <- apply(harm_final[ , 5:ncol(harm_final)], 2, function (x) { cor(x, harm_final$harmMFQ)})

harmcut <- mean(harm_cor) + 2*sd(harm_cor)
harm_words_reduce <- names(harm_cor[abs(harm_cor) >= harmcut])
```

```{r fair-analysi-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q29 <- as.character(noout1.1$Q29)

# stem the data
for (i in 1:nrow(noout1.1)) {
  noout1.1$Q29[i] <- paste(unlist(
    text_tokens(noout1.1$Q29[i], stemmer = "en")), collapse = " ")
}

##create a corpus
fair_corpus <- Corpus(VectorSource(noout1.1$Q29))
fair_TDM <- as.matrix(TermDocumentMatrix(fair_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

##view the most frequent words
fair_freq <- data.frame(Word = rownames(fair_TDM),
                       Freq = rowSums(fair_TDM),
                       row.names = NULL)
fair_freq$Word <- as.character(fair_freq$Word)
fair_freq$percent <- fair_freq$Freq/nrow(noout1.1) *100

write.csv(fair_freq, "../../study1_data/study1.1_fair_freq.csv", row.names = F)

# subset out to only words with 1% mentions
fair_words <- fair_freq$Word[fair_freq$percent >=1]
# turn sideways to add to dataset
fair_TDM <- as.data.frame(t(fair_TDM))
fair_TDM <- fair_TDM[ , fair_words]

# final dataset for fair
fair_final <- cbind(noout1.1[ , c("ResponseId", "Q15_1", "Q23", "fairMFQ")],
                   fair_TDM)

fair_cor <- apply(fair_final[ , 5:ncol(fair_final)], 2, function (x) { cor(x, fair_final$fairMFQ)})

faircut <- mean(fair_cor) + 2*sd(fair_cor)
fair_words_reduce <- names(fair_cor[abs(fair_cor) >= faircut])
```

```{r ingroup-analysis-3.1, include=FALSE}
#q27 through 34 are the writing columns
noout1.1$Q30 <- as.character(noout1.1$Q30)

##stem the data
for (i in 1:nrow(noout1.1)) {
  noout1.1$Q30[i] <- paste(unlist(
    text_tokens(noout1.1$Q30[i], stemmer = "en")), collapse = " ")
}

##create a corpus
ingroup_corpus <- Corpus(VectorSource(noout1.1$Q30))
ingroup_TDM <- as.matrix(TermDocumentMatrix(ingroup_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

# view the most frequent words
ingroup_freq <- data.frame(Word = rownames(ingroup_TDM),
                       Freq = rowSums(ingroup_TDM),
                       row.names = NULL)
ingroup_freq$Word <- as.character(ingroup_freq$Word)
ingroup_freq$percent <- ingroup_freq$Freq/nrow(noout1.1) *100

write.csv(ingroup_freq, "../../study1_data/study1.1_ingroup_freq.csv", row.names = F)

# subset out to only words with 1% mentions
ingroup_words <- ingroup_freq$Word[ingroup_freq$percent >=1]
# turn sideways to add to dataset
ingroup_TDM <- as.data.frame(t(ingroup_TDM))
ingroup_TDM <- ingroup_TDM[ , ingroup_words]

# final dataset for ingroup
ingroup_final <- cbind(noout1.1[ , c("ResponseId", "Q15_1", "Q23", "ingroupMFQ")],
                   ingroup_TDM)

ingroup_cor <- apply(ingroup_final[ , 5:ncol(ingroup_final)], 2, function (x) { cor(x, ingroup_final$ingroupMFQ)})

ingroupcut <- mean(ingroup_cor) + 2*sd(ingroup_cor)
ingroup_words_reduce <- names(ingroup_cor[abs(ingroup_cor) >= ingroupcut])
```

```{r authority-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q32 <- as.character(noout1.1$Q32)

##stem the data
for (i in 1:nrow(noout1.1)) {
  noout1.1$Q32[i] <- paste(unlist(
    text_tokens(noout1.1$Q32[i], stemmer = "en")), collapse = " ")
}

# create a corpus
authority_corpus <- Corpus(VectorSource(noout1.1$Q32))
authority_TDM <- as.matrix(TermDocumentMatrix(authority_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

# view the most frequent words
authority_freq <- data.frame(Word = rownames(authority_TDM),
                       Freq = rowSums(authority_TDM),
                       row.names = NULL)
authority_freq$Word <- as.character(authority_freq$Word)
authority_freq$percent <- authority_freq$Freq/nrow(noout1.1) *100

write.csv(authority_freq, "../../study1_data/study1.1_authority_freq.csv", row.names = F)

# subset out to only words with 1% mentions
authority_words <- authority_freq$Word[authority_freq$percent >=1]
# turn sideways to add to dataset
authority_TDM <- as.data.frame(t(authority_TDM))
authority_TDM <- authority_TDM[ , authority_words]

# final dataset for authority
authority_final <- cbind(noout1.1[ , c("ResponseId", "Q15_1", "Q23", "authorityMFQ")],
                   authority_TDM)

authority_cor <- apply(authority_final[ , 5:ncol(authority_final)], 2, function (x) { cor(x, authority_final$authorityMFQ)})

authoritycut <- mean(authority_cor) + 2*sd(authority_cor)
authority_words_reduce <- names(authority_cor[abs(authority_cor) >= authoritycut])
```

```{r purity-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q34 <- as.character(noout1.1$Q34)

# stem the data
for (i in 1:nrow(noout1.1)) {
  noout1.1$Q34[i] <- paste(unlist(
    text_tokens(noout1.1$Q34[i], stemmer = "en")), collapse = " ")
}

# create a corpus
purity_corpus <- Corpus(VectorSource(noout1.1$Q34))
purity_TDM <- as.matrix(TermDocumentMatrix(purity_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

# view the most frequent words
purity_freq <- data.frame(Word = rownames(purity_TDM),
                       Freq = rowSums(purity_TDM),
                       row.names = NULL)
purity_freq$Word <- as.character(purity_freq$Word)
purity_freq$percent <- purity_freq$Freq/nrow(noout1.1) *100

write.csv(purity_freq, "../../study1_data/study1.1_purity_freq.csv", row.names = F)

# subset out to only words with 1% mentions
purity_words <- purity_freq$Word[purity_freq$percent >=1]
# turn sideways to add to dataset
purity_TDM <- as.data.frame(t(purity_TDM))
purity_TDM <- purity_TDM[ , purity_words]

##final dataset for purity
purity_final <- cbind(noout1.1[ , c("ResponseId", "Q15_1", "Q23", "purityMFQ")],
                   purity_TDM)

purity_cor <- apply(purity_final[ , 5:ncol(purity_final)], 2, function (x) { cor(x, purity_final$purityMFQ)})

puritycut <- mean(purity_cor) + 2*sd(purity_cor)
purity_words_reduce <- names(purity_cor[abs(purity_cor) >= puritycut])
```

This procedure generated a large frequency of words for a new dictionary
of moral foundations: harm `r length(harm_words)`, fairness
`r length(fair_words)`, ingroup `r length(ingroup_words)`, authority
`r length(authority_words)`, and purity `r length(purity_words)`. These
concepts were included in the full dictionary used for Study 1.3. We
additionally created a reduced dictionary that included only concepts
correlated with their respective moral foundations scores. Correlations
between word frequency and the sum of the MFQ were calculated for each
foundation and set of concepts. Words were included in the reduced
dictionary if their correlation was two standard deviations away from
the mean correlation for that foundation. The reduced dataset included
the following number of words for each foundation: harm
`r length(harm_words_reduce)`, fairness `r length(fair_words_reduce)`,
ingroup `r length(ingroup_words_reduce)`, authority
`r length(authority_words_reduce)`, and purity
`r length(purity_words_reduce)`.

## Study 1.2

this section is about the writing

### Participants

```{r demos-mtmm-study-3.2, include=FALSE, warning=FALSE}
sample1 <- read.csv("study1_data/study1.2_fixed_sample 1.csv", stringsAsFactors = F)
sample2 <- read.csv("study1_data/study1.2_fixed_sample 2.csv", stringsAsFactors = F)

##merge files together
study1.2DF <- merge(sample1, sample2, all = TRUE)

##merge together the writing samples
#q24, q25, q26, q7 
fullsample$writing <- apply(fullsample[, c("Q24", "Q25", "Q26", "Q7")], 
                           1, function(x) toString(na.omit(x)))

##clear out completely columns we don't need
crapcolumns <- c(1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 51:61)
finalsample <- fullsample[ , -crapcolumns]

##accuracy
summary(finalsample[ , 2:40])

finalsample$Q10 <- factor(finalsample$Q10,
                         levels = 1:2,
                         labels = c("Man", "Woman"))
finalsample$Q11 <- as.numeric(finalsample$Q11)
finalsample$Q12 <- factor(finalsample$Q12,
                         levels = 1:7,
                         labels = c("White", "Black", 
                                    "Hispanic", "Asian",
                                    "Native American", 
                                    "Mixed", "Other"))

gender <- table(finalsample$Q10)/sum(table(finalsample$Q10))*100
race <- table(finalsample$Q12)/sum(table(finalsample$Q12))*100

write.csv(finalsample, 'study1_data/study1.2_finalsample.csv', row.names = FALSE)
newMFD <- import('study1_data/study1.2_finalsample_newMFD.csv')
colnames(newMFD)[1:41] <- colnames(finalsample)[1:41]
newMFD <- newMFD[,c(1,42:55)]
finalsample <- merge(finalsample, newMFD, by = "ResponseId")

>>>>

---
title: "appendix.Rmd"
---

# Study 2: Descriptive Statistics
